---
title: "Rappy Parse Run"
format: html
editor: visual
---

## Notebook to run Rappi Financial Statements

Process:

1.  Download data from google drive into the rappi prod folder
2.  For each Pdf:
    1.  apply pdf_text
    2.  extract seccion de Movimientos by capturing content between two string markers
    3.  remove unneccesary content
    4.  string split by line jump
    5.  trim
    6.  remove empty rows
    7.  ? extract gasto en periodo
    8.  Split each row into columns based on whitespace
    9.  apply str_rappy_df function to make data.frame with 3 columns: fechas, montos, conceptos
3.  Upload Results to Google Sheets

```{r setup}
pacman::p_load(
  googledrive,
  googlesheets4,
  dplyr,
  tidyr,
  stringr,
  purrr,
  tibble,
  pdftools,
  glue
)

here::i_am("Rappi Run.qmd")
```

## Download data from Google Drive

```{r driveauth}
googledrive::drive_auth(email="daniel.amieva@dar4datascience.com")
```

```{r find-pdfs}
#| echo: false
# list files 2 dowload
# estados de cuenta/ rappi credit card
files_2_download <- googledrive::drive_ls("https://drive.google.com/drive/u/0/folders/1Det2HALrcaiR1tPQMYgkjRXEIyPkgZBS",
                                         type = 'pdf' )

download_location <- here::here("inputs/rappi prod/")

```

```{r download-data}
download_rappi_statements <- function(input_path, output_path){
  googledrive::drive_download(input_path,
                              output_path,
                               "application/pdf",
                              overwrite = TRUE)
}

safely_download_rappi_statements <- safely(download_rappi_statements)

#output paths 
output_paths_pdfs <- files_2_download$name %>% 
  purrr::map_chr(
    ~glue(download_location, .x)
  )

purrr::walk2(files_2_download$id,output_paths_pdfs,
            ~safely_download_rappi_statements(.x, .y))
```

## For each pdf

```{r test-run}
rappy_test <- file.path(
  here::here('inputs/rappi prod', '202208-rappicard.pdf')
)

```

### Apply pdf_text

```{r}
# pdf_text uses lobpoopler
rappy_test_text <- pdf_text(rappy_test)
```

### Extract seccion de Movimientos by capturing content between two string markers

```{r}
start_marker <- "\\n\\n\\n\\n\\n     Movimientos"
end_marker <- "( TC\\* Tasa de conversión\\n\\n\\n\\n\\n| Compras diferidas\n\n)"

# Find the starting index of the desired text
start_index <- str_locate(text, start_marker)[, "start"]

# Find the ending index of the desired text
end_index <- str_locate(text, end_marker)[, "start"]

# Check if the start and end markers are found
if (!is.na(start_index) && !is.na(end_index)) {
  # Adjust the start index to exclude the marker
  start_index <- start_index + nchar(start_marker)
  
  # Extract the desired text using str_sub
  extracted_text <- str_sub(text, start_index, end_index - 1)
  
  # Print the extracted text
  cat(extracted_text)
} else {
  # Display a message if the markers are not found
  cat("Markers not found in the input string.")
}

```

### Remove unneccesary content

```{r}
semi_clean_text <- extracted_text %>% 
  remove_strings_with_pattern() %>% 
  str_remove_all(., "Extranjero\\n") %>% 
  #remove dollars
  str_remove_all(., "USD \\d+.\\d+") %>% 
  # remove zeors
  str_remove_all(., "\\$0\\.00")


semi_clean_text
```

### String split by line jump

```{r}
# Split the string into rows based on line jumps
rows <- strsplit(semi_clean_text, "\n")[[1]]

```

### Trim

```{r}
# Remove leading and trailing whitespace from each row
rows <- trimws(rows)
```

### Remove empty rows

```{r}
# Remove empty rows
rows <- rows[rows != ""]
```

### Split each row into columns based on whitespace

```{r}
# Split each row into columns based on whitespace
columns <- strsplit(rows, "\\s+") 
```

### Define function to remove content

```{r}
#special remove function
remove_unnecessary_chars <- function(input_string) {
  # Remove "$" from the string
  cleaned_string <- str_replace_all(input_string, "\\$", "")
  
  # Remove "," from the string
  cleaned_string <- str_replace_all(cleaned_string, ",", "")
  
  # Remove "." from the string
  cleaned_string <- str_replace_all(cleaned_string, "\\.", "")
  
  return(cleaned_string)
}
```

### Extract gasto en periodo

```{r}
#get gasto in periodo
gasto_en_periodo <- purrr::pluck(columns, -1)
clean_gasto_en_periodo <- gasto_en_periodo %>% 
  dplyr::last() %>% 
  # clean string
  remove_unnecessary_chars() %>% 
  as.numeric()
```

### Remove first and last records

```{r}
# remove last recrod and first record and third cause i establish columns
penalast_record <- length(columns) - 1
neccesary_columns <- columns[3:penalast_record]

```

### Apply str_rappy_df function to make data.frame with 3 columns: fechas, montos, conceptos

```{r}

str_rappy_df <- function(columns_string){
  fechas <- columns_string %>% 
    purrr::pluck(., 1, 1)
  
  montos <- columns_string %>% 
    purrr::pluck(., -1, 1) %>% 
    # clean string
    remove_unnecessary_chars()
  
  conceptos <- columns_string %>% 
    glue::glue_collapse(., sep = "") %>% 
    str_remove(., "\\d{4}-\\d{2}-\\d{2}") %>% 
    str_remove("\\$\\d+.\\d+") %>% 
    # clean string
    remove_unnecessary_chars() 
  
  
  print(fechas)
  print(montos)
  print(conceptos)
  
  df <- tibble(
    "fecha" = fechas,
    "monto" = montos,
    "concepto" = conceptos
  )
  
}

```

```{r}

# Create a dataframe from the columns
records_rappy <- neccesary_columns %>% 
  purrr::map_dfr(
    ~str_rappy_df(.x)
  )

# Rename the columns
colnames(records_rappy) <- c("Fecha", "Comercio", "Monto")
```

### Final Clean

```{r}
df_clean_rappy <- records_rappy %>% 
  mutate(
    Comercio = str_remove(Comercio, "\\$") %>% 
      str_remove(., ","),
    Fecha = Fecha %>% 
      lubridate::parse_date_time(., "ymd")
  ) %>% 
  tidyr::fill(Fecha, .direction = "updown")
```

## Upload Results to Google Sheets

```{r connect2sheet}
googlesheets4::gs4_auth(email = "daniel.amieva@dar4datascience.com")

rappi_statement_records_sheet <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1przfdI8MDLWf9rYPgOY9T0XLzEbhb9-xM1Gs-K6dgnY/edit#gid=1249355448",
                                                           sheet = "Finanzas Rappi Credit")
```

```{r write-results-append}
googlesheets4::sheet_append("https://docs.google.com/spreadsheets/d/1przfdI8MDLWf9rYPgOY9T0XLzEbhb9-xM1Gs-K6dgnY/edit#gid=1249355448",
                                                     df_clean_rappy,
                                                           sheet = "Finanzas Rappi Credit")
```

### Check it out by reading in

```{r checkresults}
rappy_results_test <- googlesheets4::sheet_appendgooglesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1przfdI8MDLWf9rYPgOY9T0XLzEbhb9-xM1Gs-K6dgnY/edit#gid=1249355448",
                                                     df_clean_rappy,
                                                           sheet = "Finanzas Rappi Credit")
```
